---
title: "Climate Feature Engineering"
author: "Will MacKenzie"
date: "2023-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script compares the climate attributes between WNA_BGCs. Graphical look at variables, transform of variables, and feature engineering to develop the best set of variables for mapping and future forecasting of BGC climate envelopes

Import data
```{r}

```


### 2-variable scatter plots to compare BGC averages of broader climate groups.
Start by looking at the thresholds and types from Klinka et al 2000
```{r import tables, echo=FALSE, message = FALSE}
###Climate data 1961-90for representative points per BGC  with all variables
#allDat <- fread("./inputs/ALLBGC_AllVars.csv", data.table = FALSE, stringsAsFactors = FALSE)
allDat <- fread("D:/CommonTables/HexGrids/WNA_4k_HexPts_BGCv12_Normal_1961_1990MSY_reduced6Sept2021.csv",  stringsAsFactors = FALSE,data.table = FALSE) %>% mutate(zone = 
stringr::str_extract(BGC, "\\b[A-Z]+")) %>% dplyr::select(-Tave_sm, -Tave_sp,-Tave_at,-Tave_wt)

climate <- allDat %>%  
  mutate(temp_ten = rowSums(across(starts_with("Tave"), ~ .x>10)))%>% 
  mutate(temp_zero = rowSums(across(starts_with("Tave"), ~ .x<0)))%>%  dplyr::select(BGC, temp_ten, temp_zero, MWMT, MCMT) %>% group_by(BGC)  %>%   mutate(MWMT= mean(MWMT), MCMT = mean(MCMT), na.rm = TRUE, keep = T) %>% 
  dplyr::select(BGC, temp_ten, temp_zero, MCMT, MWMT) %>% distinct(BGC, .keep_all=T)

climate.grps <- fread("./inputs/WNAv12_SubZone_Groups.csv")
climate <- left_join(climate,climate.grps)
climate2 <- climate #%>% filter(group %in% c("temperate", "temperate-warm"))
                                                                                                             
ggplot(climate2, aes(x=MCMT, y=MWMT, colour = group))+
  geom_jitter()+
  scale_color_brewer(palette = "RdYlBu")+
  geom_text_repel(size = 3, aes(label = BGC), max.overlaps = 20)+
  geom_hline(yintercept = 10)+
  geom_hline(yintercept = 20)+ 
   geom_hline(yintercept = 12)+  
geom_vline(xintercept = 0)+
  geom_vline(xintercept = -3)+
    geom_segment(x= -10 , xend = -10, y=12 , yend = 17.5, colour = "black")
```summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
