---
title: "BGC_climate_clusteranalysis"
author: "Will MacKenzie"
date: "15/10/2021"
output: html_document
---

---
title: "Random Forest Model of Biogeoclimatic Units for Western North America"
author: "William H MacKenzie & Kiri Daust"
date: "22/03/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
require (smotefamily)
require(data.table)
library(randtoolbox)
library(clhs)
library(foreign)
library(ggplot2)
library(raster)
library(RStoolbox)
library(maptools)
library(sp)
library(spatstat)
require(doParallel)
require(scales)
require(sf)
require(caret)
require (gstat)
require(ranger)
require(rticles)
#require(expss)
require(tictoc)
require(randomForest)
library(tidyverse)
require(tidymodels)
require(vip)
require(stringr)
require(corpcor)
library(yaml)
require(tidypredict)
require(skimr)
require(DescTools)
require(labdsv)
require(factoextra)
#require (gstat)
#install.packages("DescTools", dependencies = TRUE)
#data.dir = "./inputs/"
#dir.create("cLHS")

#tidymodels_prefer()
#install.packages ("spThin")
#conflicted::conflict_prefer(name = "spec", winner = "yardstick")
#conflict_prefer("rescale", "scales")

source("../Build_USA_BEC/_functions/AddVars.R")
source("../Build_USA_BEC/_functions/removeOutlier.R")
source("../Build_USA_BEC/_functions/acc_metrix.R")
cloud_dir <- "F:/OneDrive - Government of BC/CCISSv12/latest_CCISS_tool_files/"

```

Create different model variable sets
v1=all
v2 = no months
v3 = Biological Variables
vs5 = 16var
vs8 = 35 var
vs9 = reduced 35 for biologial variable reduction to 19
vs10+ = testing effects and final set

```{r create variable sets}
modelvars <- colnames(X0)[-1]
vs1 <- modelvars
month <- c("01", "02", "03", "04", "05", "06","07", "08", "09", "10", "11", "12", "Rad", "RH") ##all vars
vs2 <- modelvars[-grep(paste(month, collapse = "|"),modelvars)]
badvars <- "TD"
vs2 <- vs2[!vs2 %in% badvars] ## no monthly vars and others by choice
vs3 <- c("MAT","MCMT","MWMT","TD","MAP","MSP","AHM","SHM","DD_0","DD5","NFFD","bFFP","eFFP",
         "PAS","EMT","Eref","CMD","Tave_wt","Tave_sm","PPT_wt","PPT_sm")## biological variables from databasin
#vs5 <- read.csv("./outputs/Model_16Vars.csv") ### from original model
#vs5 <- vs5$x
vs8 <- c("Tave_at","MWMT", "DD5_at",   "DD5_sm",   "DD5_sp", "DD5_wt", # 35 variable from paper
            "Eref_sm",   "Eref_sp",
         "Tmin_wt", "MCMT","Tmin_sp",
         "PPT_JAS",  "PPT_MJ",  "PPT_sm","MSP","PPT_sp",
         "CMD_sp","CMD_at", "CMD.def", "CMDMax","CMD","CMD.total","SHM","AHM",
          "Tmax_sp", "Tmax_sm", 
          "bFFP",  "eFFP",  "NFFD", 
          "PAS_wt",  "PAS_at",  "PAS_sp", "PPT_wt","PPT.dormant","PAS_sm")


vs_final <- c("DD5", "DD_delayed",   #"DD18",#"DD5_wt","Tave_at","DD5_at", ### reduce list to omit non-biological and redundant variables
             #"Eref_sp", #"TD",
         # "EMT_threshold", #"MCMT",# EMT",  #"Tmin_sp",  #Tmin_wt",
          "PPT_MJ", "PPT_JAS", # PPT_MJ", "PPT_JAS", #  "PPT_sp","PPT_sm","MSP",
         "CMD.total", "CMI", "CMDMax", #"CMD","CMD_at",,"CMDMax"
         "SHM", "AHM", # "CMD.def",
              #"EXT",#"Tmax_sp","Tmax_sm",
            "NFFD", #"FFP",  "NFFD","NFFD_sp",
          "PAS")

```

```{r load and prep training data}
#X0 <- fread("./inputs/training_data/WNAv12_2km_Hex_Pts_Normal_1961_1990MSY_3.csv")
X0 <- fread("D:/CommonTables/HexGrids/WNA_4k_HexPts_BGCv12_Normal_1961_1990MSY_reduced6Sept2021.csv") %>% dplyr::filter(!(BGC == "MHmm2" & Latitude > 55)) ##ClimateNA data
IDVars <- X0[, .(ID1, Latitude, Longitude, Elevation)]

```

```{r reduce variables}
X0 <- addVars(X0)
X0 <-
  X0[, c("ID1", "BGC", vs_final), with = F] 

## select the variable set here
count_tp <- X0 %>% dplyr::count(BGC)
#X0 <- X0 %>% filter(!BGC == "ESSFab")
```

The preprocessing function from the caret package was used to identify variables with near-zero variance or correlation >0.90 in the combined data set. These variables were removed leaving a final variable set of 20 variables.


```{r correlated, warning=FALSE} 
  # correlationMatrix <- cor(X0[,-c(1:2)])
  # highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.7, verbose = TRUE, names = TRUE) ## review for removal of highly correlated vars
  #  highlyCorrelated
  # X0 <- X0 %>% dplyr::select(-c(highlyCorrelated))%>% drop_na(BGC)

```
```{r summarize climates}

BGC.clim <- X0 %>% dplyr::select(-ID1) %>% group_by(BGC) %>% summarize_all(mean)

```
# 1. Comparison of predicted ratio of BGC subzone/variants from all GCM models
```{r create data matrix, include=FALSE}
dat.long <- BGC.clim %>% pivot_longer(!BGC, names_to = 'climvar', values_to = 'value')  %>% data.table
dat.long2 <- dat.long %>%  filter(str_detect(BGC, "IDF"))

dat.matrx <- matrify(dat.long) %>% decostand(method = 'standardize')

```


### 1a. Optimal cluster groups 

``` {r cluster analysis all, echo = F, result = F, messages = FALSE, warning = 'hide'}
n_clust <- fviz_nbclust(dat.matrx, kmeans, method = "silhouette")#+k.max = 8,
plot(n_clust) ## "optimal" cluster number
n_clust<-n_clust$data
best.clust<-as.numeric(n_clust$clusters[which.max(n_clust$y)])
```

### 1b. Cluster Analysis of 2050s rcp45 BGC predictions by GCM (all GCMs)

```{r output dendrogram, warning=FALSE}
best.clust <- 4
gcm_clst <- hcut(dat.matrx, k = best.clust, hc_method = "ward.D2", hc_metric = "euclidean")
gcm.group <- gcm_clst$cluster %>% as.data.frame %>% rownames_to_column("BGC")
gcm_dend <- fviz_dend (gcm_clst ,  color_labels_by_k = T, lwd = .5,
                       rect = T, cex = .5, horiz = T, scale = "none") %>% 
  plot()
#ggsave("./outputs/BGC_ClusterDendrogram_2km_euclid.jpeg")


# fviz_dend(gcm_clst, cex = 0.5, lwd = 0.8, k = 2, 
#           rect = TRUE, 
#           k_colors = "jco", 
#           rect_border = "jco", 
#           rect_fill = TRUE,
#           type = "phylogenic",
#           repel = TRUE,
#           phylo_layout = "layout_as_tree")


```

### 1c. Principle Components with GCM clusters (all GCMs)

```{r output PCO}
# Visualize
fviz_cluster(gcm_clst, data = dat.matrx,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal(),
             main = "PCO of BGC clusters")


fviz_pca_biplot(gg.pca, label = "var", habillage=gcm_clst$cluster,
               addEllipses=TRUE, ellipse.level=0.95,
               ggtheme = theme_minimal())

## clusterability###numbers below 0.5 have low clusterability
res <- get_clust_tendency(dat.matrx, n = nrow(dat.matrx)-1, graph = F)
res$hopkins_stat
```
### PCA biplot of selected environmental variables

```{r PCA of env, fig.cap = "PCA biplot of selected environmental variables by Site Series"}
#env <- Env1 %>% select(MinCN, Light, N, SMRmean, Mg,  OrganicP, S, MinpH)
#env <- Env1 %>% select(-SS, -Transect, -SiteSeries, -Lat, -Long, -Elev, -BasalA, -TreeHt) %>% select(-starts_with('Pinus') , -starts_with('Abies'))
require(vegan)
require(ggbiplot)
require(hrbrthemes)
  env.std <- decostand(dat.matrx, method = "standardize")
pca.env <- rda(env.std)
  env_pca <- ade4::dudi.pca(env.std, nf = 5, scannf = F)# sqrt to create euclidean space

  gg.pca <- prcomp(env.std, scale = TRUE)
    #groups <- as.factor(Env1$SS)
  
  env_biplot <- ggbiplot(gg.pca,obs.scale = 1, var.scale = 1,
    groups = as.factor(gcm.group$.), ellipse = TRUE, circle = FALSE) +
  #scale_color_discrete(name = '') +
  #theme(legend.direction = 'horizontal', legend.position = 'top')+
  scale_fill_ipsum() #+
    #theme_ipsum()
  plot(env_biplot)
  #ggsave(env_biplot, filename = "./figures/PCA.png", device = "png", width = 6, height = 8)
#   groups <- as.factor(Env1$SiteSeries)
# pca_plot <- fviz_pca_biplot(env_pca,
#                             geom.ind = c("point"),
#              col.ind = groups, # color by groups
#              #palette = c("#00AFBB",  "#FC4E07"),
#              addEllipses = TRUE, # Concentration ellipses
#              ellipse.type = "confidence",
#              legend.title = "Site Series",
#              repel = TRUE,
#              title = "")
# plot(pca_plot)


```
