---
title: "LocalFeatureSelection"
author: "KDaust and WHMacKenzie"
format: html
editor: visual
---

## Focal BGC comparison to neighbours

This code builds randomforest models of every BGC in WNA compared to spatial neighbours and returns the model error and variables with high importance in each model. The aim of this approach is an alternate way to create a list of variables to include in the WNA BGC model.


```{r package and functions}
library(data.table)
library(ranger)
library(foreach)
library(terra)
library(sf)
library(dplyr)
library(climr)
library(vip)
library(tictoc)
require(ggplot2)
require(stringr)

addVars <- function(dat){ ##this function modifies everything inplace, so no need for return value
  dat[,`:=`(PPT_MJ = PPT05+PPT06,
            PPT_JAS = PPT07+PPT08+PPT09,
            PPT.dormant = PPT_at+PPT_wt)]
  dat[,`:=`(CMD.def = 500-PPT.dormant)]
  dat[CMD.def < 0, CMD.def := 0]
  dat[,`:=`(CMDMax = CMD07,
            CMD.total = CMD.def + CMD)]
  dat[,`:=`(CMD.grow = CMD05+CMD06+CMD07+CMD08+CMD09,
            DD5.grow = DD5_05+DD5_06+DD5_07+DD5_08+DD5_09,
            #DDgood = DD5 - DD18,
            #DDnew = (DD5_05+DD5_06+DD5_07+DD5_08)-(DD18_05+DD18_06+DD18_07+DD18_08),
            TmaxJuly = Tmax07)]
}

```

This chunk needs to be run only once for a version of BGC maps

```{r build comparative list}
# bgc_map <- st_read("D:/CommonTables/CCISS/WNA_BGC_v12_19Apr2024.gpkg") %>% dplyr::filter(!BGC == "(None)")
# bgc_map$ID <- seq_along(bgc_map$BGC)
#  bgcs <- unique(bgc_map$BGC) %>% data.frame
#  fwrite(bgcs, "wna_bgcs.csv")
#  
#  bgc_info <- fread("D:/CommonTables/CCISS/WNA_BGCs_Info_v12_16.csv") %>% filter(BGC %in% bgcs$.) #%>% mutate(bgc = as.factor(BGC))
# BC_BGCs <- bgc_info[grep("BGC.*",Source),BGC]
# 
# #bgc_map <- st_read("BC_BGCs_with_ID.gpkg")
# 
# neighbours_ls <- list()
# #for(bgc in BC_BGCs[-(1:180)]){
# for(bgc in bgc_info$BGC){
#   cat(".")
#   focal <- bgc_map[bgc_map$BGC == bgc,]
#   if(nrow(focal) > 0){
#     focal <- st_union(focal$geom)
#     neighbours <- st_intersects(focal, bgc_map)
#     neighbours_ls[[bgc]] <- neighbours[[1]]
#   }
# }
# 
# saveRDS(neighbours_ls,"WNA_BGC_NeighbourList.rds",)
# 
# st_write(bgc_map, "WNA_BGCs_with_ID.gpkg", append = FALSE)
# 
dem <- terra::rast("D:/CommonTables/DEMs/WNA_DEM_4326_clipped.tif")
# saveRDS(dem, "//objectstore2.nrs.bcgov/ffec/CCISS_Working/WNA_DEM_4326_clippedDEM.rds")
```
### Build randomforest models
This chunk loops through the comparative list from previous chunk and creates a ranger model for each focal BGC and its neighbours.
The model OBO and top variables is returned from each iteration as base working data for summarizing.
This run can take considerable time.

```{r build random forests}
tic()
neighbours_ls <- readRDS("WNA_BGC_NeighbourList.rds")
remove1 = c("PPT_at", "PPT_wt", "CMD.def")
remove2 = c("Eref_at", "PPT.dormant", "TD", "PPT_sm", "NFFD_wt", "MSP", "MAP", "DD18_wt", "PPT", "CMD_at", ## high use by descriptive only
            'PAS_sm', 'NFFD_at', 'Tmax_at', 'DD5_sm', 'NFFD', 'DD18', 'DD_18', 'DD_18_sm', 'MAT', 'Tave', 'DD5', 
            'DD_18_at', 'DD5.grow', 'DD18_at', 'DD5_at')
remove3 = c(FFP, Tmax, Tave_at, eFFP, bFFP, DD_0, #less thatn 10 uses)
            MCMT, EMT,PAS_at, PPT_sp, DD_0_wt, DD18_sp, DD18_wt) 
vars.selected <- fread("no_month_vars.csv") %>% 
  dplyr::filter(!vars %in% remove1) %>% 
    dplyr::filter(!vars %in% remove2)

bgc_list <- names(neighbours_ls)
bgc_list <- bgc_list[!grepl("BAFA.* | *.un.", bgc_list)]
bgc_list <- bgc_list[-1]

tic()

vars <- c(vars.selected$vars,"BGC")
bgc_list_short <- c("BGxh3", "BWBSdk", "CDFmm", "ICHmc1", "CWHmm1", "ESSFwk1","ICHmw1","IDFdk3",
                   "IDFxx2","MSdw","SBSdk","MSxv")
#bgc_list <- bgc_list[-"CWHvm2"]
tic()
res_list <- list()
#bgc = "CDFmm"
for(bgc in bgc_list){
  cat("Processing",bgc,"\n")
  out <- bgc_map[bgc_map$ID %in% neighbours_ls[[bgc]],]
  out_union <- group_by(out, BGC) %>% 
    summarize(geom = st_union(geom),
              BGC = BGC[1])
  pnts <- st_sample(out_union, size = rep(150, nrow(out_union)), type = "random", by_polygon = T)
  pnts_all <- st_as_sf(data.frame(BGC = rep(out_union$BGC, each = 150), geometry = pnts))
  pnts_all <- st_transform(pnts_all, 4326)
  coords <- st_coordinates(pnts_all)
  temp_elev <- terra::extract(dem, coords)
  coords.bgc <- data.frame(coords, elev = temp_elev$WNA_DEM_SRT_30m,
                       ID = 1:nrow(pnts_all), BGC = pnts_all$BGC) %>% rename(lat = Y, lon = X, id = ID) %>% 
    dplyr::select(lon, lat, elev,id, BGC)
  coords <- coords.bgc %>%   dplyr::select(lon, lat, elev,id)
  coords.bgc <- coords.bgc %>% dplyr::select(id, BGC)
  # coords <- data.frame(coords, elev = temp_elev$WNA_DEM_SRT_30m, 
  #                      ID = 1:nrow(pnts_all), BGC = pnts_all$BGC)
  
  clim_vars <- suppressMessages(climr_downscale(coords, which_normal = "auto", 
                                                vars = list_variables(), return_normal = T))  
  clim_vars <- data.table:::na.omit.data.table(clim_vars)
  addVars(clim_vars)
  clim_vars <- left_join(clim_vars, coords.bgc)
  
  clim_vars <- setDT(clim_vars)[,..vars]
  
  clim_vars[,BGC := as.factor(BGC)]
  rf_mod <- ranger(BGC ~ ., data = clim_vars, num.trees = 101, importance = "impurity", splitrule = "gini")
  varimp <- sort(importance(rf_mod),decreasing = T)[1:6]
  res_list[[bgc]] <- data.table(Focal = bgc, Var = names(varimp), 
                                Importance = unname(varimp), 
                                OOB = rf_mod$prediction.error,
                                NumberBGCs = nrow(out_union))
}


dat_all2 <- rbindlist(res_list)
fwrite(dat_all2, "Focal_Variable_Importance_WNAv2.csv")
toc()
```

```{r summarize results}
dat_all1 <- fread("Focal_Variable_Importance_WNAv1b.csv") %>% mutate(varset = "1")
mean.error1 <- dat_all1 %>% select(Focal, OOB, varset ) %>% distinct %>%  mutate(mean.error = mean(OOB))
dat_all2 <- fread("Focal_Variable_Importance_WNAv2.csv") %>% mutate(varset = "2")
mean.error2 <- dat_all2 %>% select(Focal, OOB, varset ) %>% distinct %>%  mutate(mean.error = mean(OOB))
mean.error <- rbind(mean.error1, mean.error2)


mean.error2 <- mean.error %>% filter(!str_detect(Focal, "p$"))%>% filter(!str_detect(Focal, "w$"))
require(ggplot2);require(stringr)
ggplot(mean.error, aes(x= varset , y=OOB, fill = varset))+
  geom_violin()+
  stat_summary(fun.data = "mean_sdl",  geom="crossbar", width=0.05)

var.count <- dat_all2[,.(Num = .N), by = .(Var)]
setorder(var.count, -Num)

fwrite(var.count, "Count_of_Variables_v2.csv")

var.type <- fread("Variable_Types.csv")
dat_all2 <- fread("Focal_Variable_Importance_WNAv1.csv") %>% left_join(var.type)
BGC_using_group <- dat_all2 %>% group_by(Focal, var.group, var.season) %>% count() %>% select(-n) %>%  ungroup() %>% group_by(var.group, var.season) %>% count()
var.type.count <- dat_all2[,.(Num = .N), by = .(var.type)]
setorder(var.count, -Num)
```


```{r map of BGCs with poor OBO}
OOB <- dat_all2 %>% select(Focal, OOB) %>% rename(BGC = Focal)
bgc_map <- st_read("D:/CommonTables/CCISS/WNA_BGC_v12_19Apr2024.gpkg") %>% dplyr::filter(!BGC == "(None)") %>% left_join(OOB)
### add OBO to BGC
require(tmap)
tm_shape(bgcmap +tm_fill(fill = "OOB"))
# ggplot(data = bgc_map) +
#     geom_sf(aes(fill = OOB)) +
#     scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

