---
title: "QML_to_CSV"
author: "KaraSchaefer_WillMacKenzie"
date: "28/03/2020"
output: html_document
---



```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(ggplot2)
library(XML)
library(colortools)
library(sf)
library(lwgeom)
library(raster)
library(sp)
library(data.table)
library(dplyr)
library(bcdata)
library(bcmaps)


```
##Script to assign colours to map units in a PEM based on complementary colours.
The user will read in a Map Unit Legend and this script will assign and write colours back to the legend files.  
The zonal site series colour will match the colours used for the BGC subzone variant
Then complementary, adjacent, or analagous colours will be selected based on edatopic position
Special colours/patterns will be assigned to non-forested classes.
Some links to script descriptions: 
https://www.datanovia.com/en/blog/how-to-create-beautiful-color-palettes-in-r/

Online colour tool
http://colorschemedesigner.com/csd-3.5/


```{r import BGC units and map_keys, tidy = TRUE, warning=FALSE}
file_name <- "WNA_SubzoneColours"

document <- xmlParse("WNAv11_Subzones.qml")
category_nodes <- getNodeSet(document, "//category")
BGC_zone <- lapply(category_nodes, xmlGetAttr, "value")
xml_id <- lapply(category_nodes, xmlGetAttr, "symbol")
BGC_with_xml_id_only <- data.frame(BGC_zone = unlist(BGC_zone), xml_id = unlist(xml_id))

symbol_nodes <- getNodeSet(document, "//symbols//symbol[@name]")
symbol_id <- sapply(symbol_nodes, xmlGetAttr, "name")
names(symbol_id) <- sapply(getNodeSet(document, '//symbols//symbol//prop[@k="color"]'), xmlGetAttr, "v")

id_and_colour <- data.frame(id = symbol_id,
                            cmyk_colour = names(symbol_id), stringsAsFactors = FALSE)

BGCs_with_colours <- BGC_with_xml_id_only %>% 
  left_join(id_and_colour, by = c("xml_id" = "id"))

BGCs_with_colours <- BGCs_with_colours[, c("BGC_zone","cmyk_colour", "xml_id") ]


SUBZONE_COLOURS_KEY <- BGCs_with_colours %>% 
  mutate(colour = sapply(
    strsplit(BGCs_with_colours$cmyk_colour, ","),
    function(x) rgb(x[1], x[2], x[3], maxColorValue = 255)),
    classification = BGC_zone)

SUBZONE_COLOURS_KEY <- SUBZONE_COLOURS_KEY[, c("classification", "colour")]
write.csv (SUBZONE_COLOURS_KEY, "WNAv11_Subzone_Colours.csv")

#BGCs <- read.csv(file.path(map_unit_legend, "AllModelled_BGCs_v11.csv")) ### BGC RGB colour will be assigned here manually picked to match the BGC field maps
#BGCs <- BGCs[!is.na(BGCs),]


# alternative to use bc catalogue or another dataset with unique bgc 
#bec_sf <- bec(class = "sf")
#bec_sf <- bcdc_query_geodata("0d526b63-812b-444c-b227-418b6fcbbd44")%>%
#      bcdata::select(c("BG_CODE")) %>% # Treed sites
#      collect()

```


```{r generate .qml file, tidy = TRUE}

alpha = 255
base = newXMLNode("qgis")
addAttributes(base,
              minScale="1e+08",
              version="0.0.1",
              maxScale="0",
              hasScaleBasedVisibilityFlag="0",
              styleCategories="AllStyleCategories")
flags <- newXMLNode("flags")
identifiable <- newXMLNode("Identifiable", 1)
removable <- newXMLNode("Removable", 1)
searchable <- newXMLNode("Searchable", 1)

custom_properties <- newXMLNode("customproperties")
property_a <- newXMLNode("property", attrs = c(value="false",
                                               key="WMSBackgroundLayer"))
property_b <- newXMLNode("property", attrs = c(value="false",
                                               key="WMSPublishDataSourceUrl"))
property_c <- newXMLNode("property", attrs = c(value="0",
                                               key="embeddedWidgets/count"))
property_d <- newXMLNode("property", attrs = c(value="Value",
                                               key="identify/format"))

pipe <- newXMLNode("pipe")
raster_renderer <- newXMLNode("rasterrenderer",
                              attrs = c(band="1",
                                        type="paletted",
                                        opacity="1",
                                        alphaBand="-1"))
raster_transparency <- newXMLNode("rasterTransparency")
min_max_origin <- newXMLNode("minMaxOrigin")
limits <- newXMLNode("limits", "None")
extent <- newXMLNode("extent", "WholeRaster")
stat_accuracy <- newXMLNode("statAccuracy", "Estimated")
cumulative_cut_lower <- newXMLNode("cumulativeCutLower", 0.02)
cumulative_cut_upper <- newXMLNode("cumulativeCutUpper", 0.98)
standard_deviation_factor <- newXMLNode("stdDevFactor", 2)

colour_palette <- newXMLNode("colorPalette")
palette_entry <- lapply(seq_along(colour_coded_table$integercode),function(val){newXMLNode("paletteEntry",
                                                                                           attrs = c(value = colour_coded_table$integercode[val],
                                                                                                     alpha = alpha,
                                                                                                     label = as.character(val),
                                                                                                     color = colour_coded_table$colour[val]))
})
addChildren(colour_palette,palette_entry)

colour_ramp <- newXMLNode("colorramp", attrs = c(type="randomcolors",
                                                 name="[source]"))

brightness_contrast <- newXMLNode("brightnesscontrast", attrs = c(brightness=0,
                                                                  contrast=0))
hue_saturation <- newXMLNode("huesaturation", attrs = c(colorizeStrength=100,
                                                        saturation=0,
                                                        colorizeRed=255,
                                                        grayscaleMode=0,
                                                        colorizeGreen=128,
                                                        colorizeBlue=128,
                                                        colorizeOn=0))

blend_mode <- newXMLNode("blendMode", 0)

addChildren(flags, list(identifiable,
                        removable,
                        searchable))
addChildren(custom_properties, list(property_a,
                                    property_b,
                                    property_c,
                                    property_d))
addChildren(pipe, list(raster_renderer,
                       brightness_contrast,
                       hue_saturation))
addChildren(raster_renderer, list(raster_transparency,
                                  min_max_origin,
                                  colour_palette,
                                  colour_ramp))
addChildren(min_max_origin, list(limits,
                                 extent,
                                 stat_accuracy,
                                 cumulative_cut_lower,
                                 cumulative_cut_upper,
                                 standard_deviation_factor))
addChildren(base, list(flags,
                       custom_properties,
                       pipe,
                       blend_mode))

writeLines(saveXML(base), file.path(examples, "example.qml"))
```


